---
title: "Final Exam"
output: html_document
date: "2022-12-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions

This final exam is open book, open note, and you are welcome to use any code from class or from your previous homework assignments.  **Please do not collaborate with any other students.**  Answer the following questions in the spaces provided. For any written explanations, write answers *outside* of code chunks. Insert code chunks and write code for answers that require coding.

Knit the .Rmd file when you are finished (also knit along the way to catch bugs in code early). Make sure your code and output are visible in the knitted document. Submit *both* the knitted HTML file *and* the .Rmd file to Canvas. (You should probably check over the knitted document before submitting it.)

## Question 1

Imagine that you are in charge of data and analytics for Pete Buttigieg's 2028 presidential campaign.  The latest round of tracking polls for the upper midwest has just come across your desk, with the following results:

```{r echo = FALSE}
State = c('Wisconsin','Michigan','Minnesota','Indiana','Illinois')
Buttigieg_voters = c(595,485,131,1239,1561)
number_of_respondents = c(1312,985,251,2501,2710)
tracking_polls = data.frame(State,Buttigieg_voters,number_of_respondents)
print(tracking_polls)
```

The campaign weekly strategy meeting, in which allocations of campaign resources will be decided, is scheduled for this afternoon.  On the basis of this data, which states do you recommend increasing spending in and why?

```{r}
MOE = 1.96*sqrt((tracking_polls$Buttigieg_voters/tracking_polls$number_of_respondents * (1 - tracking_polls$Buttigieg_voters/tracking_polls$number_of_respondents))/tracking_polls$number_of_respondents)

MOE
tracking_polls$Buttigieg_voters/tracking_polls$number_of_respondents
tracking_polls$Buttigieg_voters/tracking_polls$number_of_respondents - MOE
tracking_polls$Buttigieg_voters/tracking_polls$number_of_respondents + MOE

```

## Question 2

Interpret the following regression table.

```{r}

```
